Self-tunable approximated explicit MPC: Heat exchanger implementation and analysis

### 2023-06-21 paper/ATE/main.tex
ATE update3

	\textcolor{red}{From the point of computational complexity, the proposed tuning procedure does not lead to any significantly demanding mathematical operations. Simple relations in Eq.~\eqref{eq:rho_delta_ref} and Eq.~\eqref{eq:rho_tilde_def} are evaluated.}

@@

The system was identified by experimental identification. More specifically, several step responses were obtained in the whole range of the input variable steps. Every step response was identified as a first-order system and the nominal transfer function was calculated. Finally, the transfer function was converted to the state-space model. The matrices of the discrete-time state-space model of the plant are
\begin{subequations}
	\label{eq:model_A_B} 
	\begin{eqnarray}
		A = \begin{bmatrix}
			0.839
		\end{bmatrix}, \quad
		B = \begin{bmatrix}
			0.039
		\end{bmatrix}, \quad
		C = \begin{bmatrix}
			1
		\end{bmatrix}, 
		%\quad
		%D = \begin{bmatrix}
		%0
		%\end{bmatrix}.
	\end{eqnarray}
\end{subequations}
considering the sampling time $T_\mathrm{s}$ = 1\,s. 
To respect the physical limitations of the operating conditions, the constraints are considered in the terms of control inputs
\begin{eqnarray}
	\label{eq:u_const}
	-15\% \le u \le 65\%,
\end{eqnarray}
where the variable $u$ represents the control inputs in the deviation form. The values of feed temperature and voltage of the heating medium pump corresponding to zero steady states are respectively $T^\mathrm{s}$~=~35 $^{\circ}\mathrm{C}$ and $U^\mathrm{s}$~=~35\%. Therefore, the physical constraints on the manipulated variable are actually
\begin{eqnarray}
	\label{eq:U_const}
	20\% \le U \le 100\%.
\end{eqnarray}	
As the controlled system is naturally stable even if the maximal or minimal value of the manipulated variable is constantly applied, the constraints on the controlled variable in Eq.~\eqref{eq:mpc_problem_state_constraints} can be omitted. Another motivation to include the output/state constraints would be the necessity to retain the controlled variable in some comfort range, specified by operation requirements. As it was not the aim of the case study, this reason also leads to omitting the constraints in Eq.~\eqref{eq:mpc_problem_state_constraints} to fully investigate the control performance of the pure self-tuning method.

@@

@@

	\section*{Acknowledgments}

>>>>

	\textcolor{red}{From the point of computational complexity, the proposed tuning procedure does not lead to any significantly demanding mathematical operations. Simple relations in Eq.~\eqref{eq:rho_delta_ref} and Eq.~\eqref{eq:rho_tilde_def} are evaluated. Q: Mergnut s tym, co som pridala ako dosledky tunable eMPC ako takeho? Ze sa zvysi pamat a cas, lebo sa vyuzivaju dva eMPC?}

@@

The system was identified by experimental identification. \textcolor{red}{The aim was to work with linear nominal model in MPC optimization problem to decrease the numerical complexity. To avoid plant-model mismatch in order to ensure offset-free tracking, either disturbance observer or built-in integrator (Eq.~\eqref{eq:mpc_problem}) can be employed. Note, in this work, the built-in integrator was considered. The system was identified based on several step responses. The step changes were performed in the whole range of admissible values of manipulated variable and every step response was identified by transfer function. It was possible to identify every step response as a first-order system, while the nominal gain and time constant are respectively $K = 0.2418\,^{\circ}\mathrm{C}$ and $\tau = 5.7$\,s. (TODO: checknut units pre gain a pridat symboly do nomenklatury, ak sa zhodneme na oznaceni)} 
%More specifically, several step responses were obtained in the whole range of the input variable steps. Every step response was identified as a first-order system and the nominal transfer function was calculated.
Finally, the nominal transfer function was converted to the state-space model. The matrices of the discrete-time state-space model of the plant are
\begin{subequations}
	\label{eq:model_A_B} 
	\begin{eqnarray}
		A = \begin{bmatrix}
			0.839
		\end{bmatrix}, \quad
		B = \begin{bmatrix}
			0.039
		\end{bmatrix}, \quad
		C = \begin{bmatrix}
			1
		\end{bmatrix}, 
		%\quad
		%D = \begin{bmatrix}
		%0
		%\end{bmatrix}.
	\end{eqnarray}
\end{subequations}
considering the sampling time $T_\mathrm{s}$ = 1\,s. 
To respect the physical limitations of the operating conditions, the constraints are considered in the terms of control inputs
\begin{eqnarray}
	\label{eq:u_const}
	-15\% \le u \le 65\%,
\end{eqnarray}
where the variable $u$ represents the control inputs in the deviation form. The values of feed temperature and voltage of the heating medium pump corresponding to zero steady states are respectively $T^\mathrm{s}$~=~35 $^{\circ}\mathrm{C}$ and $U^\mathrm{s}$~=~35\%. Therefore, the physical constraints on the manipulated variable are actually
\begin{eqnarray}
	\label{eq:U_const}
	20\% \le U \le 100\%.
\end{eqnarray}	
As the controlled system is naturally stable even if the maximal or minimal value of the manipulated variable is constantly applied, the constraints on the controlled variable in Eq.~\eqref{eq:mpc_problem_state_constraints} can be omitted. Another motivation to include the output/state constraints would be the necessity to retain the controlled variable in some comfort range, specified by operation requirements. As it was not the aim of the case study, this reason also leads to omitting the constraints in Eq.~\eqref{eq:mpc_problem_state_constraints} to fully investigate the control performance of the pure self-tuning method.

@@

\textcolor{red}{An interesting phenomenon can be observed during tracking the third reference value, i.e. $T_{\mathrm{ref}}$ = 50\,$^{\circ}\mathrm{C}$. Although the steady-state values of temperature have the same value, the values of the manipulated variable are different. To check the correctness of the results, the measurements were performed multiple times and led to the same behavior. The reason could be explained by the peak of the manipulated variable associated with the upper boundary controller at time 800\,s. After approximately 100\,s the value of manipulated variable dropped and settled at a value that was lower than the value associated with the lower boundary controller. This is a consequence of the heat accumulated inside the heat exchanger plates, and therefore less heating medium was necessary to heat the feed. This phenomenon does not happen when tracking the reference value $T_{\mathrm{ref}}$ = 45\,$^{\circ}\mathrm{C}$, which can be explained by the nonlinear behavior of the heat transfer process. When working in higher temperatures, the heat transfer process gain decreases and the sensitivity to changes in the heating medium flow is lower. Therefore even different flow rates of the heating medium lead to the same temperature at the outlet. }

@@

	\textcolor{red}{Note, this strategy relies on a proper design of the two boundary controllers. In case a disturbance occurs, both boundary controllers should be able to solve a disturbance rejection problem, as the final value of the manipulated variables is interpolated between them. The future work will focus on incorporating also the measurable disturbances into the tuning strategy.}
	\section*{Acknowledgments}
